---
title: "Reproducible research peer assignment 1"
author: "vjanes"
date: "Sunday, December 14, 2014"
output: html_document
---

Peer assignment one for the Coursera Data Science specialty's *Reproducible research* course.

##Loading and preprocessing the data
```{r}
Sys.setlocale("LC_TIME", "English")
library(lattice)
```

## Reading datafiles

```{r}
unzip("activity.zip")
data<-read.csv("activity.csv", colClasses = c("integer","Date","integer"))
```

##What is mean total number of steps taken per day?

### Aggregating data by day
Using aggregate function
```{r}
stp_per_day<-aggregate(data$steps, by=list(data$date), "sum",na.rm=TRUE, na.action=NULL)
names(stp_per_day)<-c("date","tt_steps")
```


### Producing an histogram
Using barplot

```{r}
barplot(stp_per_day$tt_steps, names.arg=stp_per_day$date,width = 0.1,main="Total number of steps by day")
```



### Calculating mean and median "total steps per day" values
Using common R functions

```{r}
stp_mean<-mean(stp_per_day$tt_steps)
stp_med<-median(stp_per_day$tt_steps)  
  
```

The mean number of steps per day is **`r stp_mean`**.  
The median number of steps per day is **`r stp_med`**.

##What is the average daily activity pattern?

### Aggregating data by time-period
Using aggregate function
```{r}
stp_per_tp<-aggregate(data$steps, by=list(data$interval), "mean",na.rm=TRUE, na.action=NULL)
names(stp_per_tp)<-c("interval","mean_steps")
```

### Producing an histogram

```{r}
plot(stp_per_tp$interval,stp_per_tp$mean_steps, type="l",
     xlab="time-period", ylab="mean steps",main="Mean number of steps by time-period")
```

### Max time-period
time-period with max mean steps:

```{r}
max_mean_step_tp<-
  stp_per_tp[which.max(stp_per_tp$mean_steps),1]

max_mean_step_val<-round(
  stp_per_tp[which.max(stp_per_tp$mean_steps),2],digits=2)

```

The time period with max mean number of steps is interval **`r max_mean_step_tp`**.  
The max mean steps value is **`r max_mean_step_val`**.

##Imputing missing values

### Total number of missing values in the dataset

```{r}
nb_na_rows<-NROW(data[,1])-NROW(na.omit(data[,1]))
```
There are **`r nb_na_rows`** out of `r NROW(data[,1])` rows.

###  Filling in all of the missing values
In order to fill missing value we will feed median values for interval in the dataset. In order to do this we'll create a modified datasample names _**datam**_.

```{r}
datam<-data
for (i in (1:NROW(datam)))
{
  if(is.na(datam[i,1]))
    {datam[i,1]<-
       stp_per_tp[which(stp_per_tp$interval==datam[i,3]),2]}}
```

### Producing an histogram for completed data

```{r}
stp_per_daym<-aggregate(datam$steps, by=list(datam$date), "sum",na.rm=TRUE, na.action=NULL)
names(stp_per_daym)<-c("date","tt_steps")
barplot(stp_per_daym$tt_steps, names.arg=stp_per_day$date,width = 0.1, main = "Total steps per day, 
 completed sample")

```


```{r}
stp_meanm<-mean(stp_per_daym$tt_steps)
stp_medm<-median(stp_per_daym$tt_steps)  
  
```

The mean number of steps per day is **`r stp_meanm`** for completed sample.  
The median number of steps per day is **`r stp_medm`** for completed sample.  **Both mean and median have slightly increased**.

##Are there differences in activity patterns between weekdays and weekends?

###Defining WE/non-WE days

Creating the WE/non-WE table:
```{r}
WE_table<-cbind(
  rbind("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"),rbind("non_WE","non_WE","non_WE","non_WE","non_WE","WE","WE"))
colnames(WE_table)<-c("weekday","is_we")
```
Applying to data, displaying result sample:
```{r}
dataw<-cbind(data,weekday=weekdays(data$date))
dataw<-merge(dataw,WE_table,by="weekday")

##displaying result sample
samdataw<-dataw[which(dataw$interval==0),c(3,1,5)]
samdataw<-samdataw[order(samdataw$date),]
row.names(samdataw) <- NULL 
head(samdataw[order(samdataw$date),],n=10)
```

###Comparing WE/non-WE days

```{r}
stp_interv_we<-aggregate(dataw$steps, by=list(dataw$interval,dataw$is_we), "mean",na.rm=TRUE, na.action=NULL)
colnames(stp_interv_we)<-c("interval","is_we","steps")

xyplot(steps~interval|is_we, data=stp_interv_we, type="l",layout=c(1,2))
```

